|

### 2023-07-18 paper/03-methods.Rmd
updated bootstrap and more edits

In addition to the average STRAIN over all frames played, we can also calculate pass rusher $j$'s observed average STRAIN on a single play $i$ consisting of $T_i$ total frames,
\begin{equation*}
\overline{\text{STRAIN}}_{ij} = \frac{1}{T_i} \sum_{t=1}^{T_i} \widehat{\text{STRAIN}}_{ij}(t).
\end{equation*}
While this aggregated measure is a simple first step for pass rush evaluation, the observed average STRAIN on a single play is likely due to a convolution of factors.
Outside of the pass rusher's ability, there is variability in the opposing strength of pass blockers across plays a pass rusher is involved - both at the individual and team level.
Additionally, based on previous work by PFF, the opposing QB plays an important role on the observed pressure in passing plays [@eager2019pressure]. We also suspect that pass rusher's STRAIN is affected by their defensive teammates as well. 
Thus, we need to appropriately divide the credit of an observed average STRAIN across the different players and team involved, amongst other factors.

@@

We consider a normal distribution to shrink the random intercepts for each player and team toward their respective group means.
This is a useful property since we do not observe the same number of plays for each player. 
For the team effects, this provides us with the average defense and offense team-level effects on a pass rusher's STRAIN.
Due to the nested nature of players on teams, the individual pass rusher and blocker random intercepts reflect the respective player's effect relative to their team effects. 
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

@@

As for the fixed-effects about pass rusher $j$ in play $i$, we include a variety of features that likely contribute to variation in STRAIN.
First, we adjust for the position of both the pass rusher and nearest blocker to account for any positional effects.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
These are encoded as indicator variables with defensive ends and tackles as the reference levels for the pass rushers and blockers, respectively.
We also adjust for the number of pass blockers on the play, since teams may decide to employ a more protective scheme that could lower the observed STRAIN.
Finally, we control for play-context covariates with respect to the offensive team.
These include the current down (first, second, third, fourth, or two-point conversion), yards to go for a first down, and current yardline (i.e., distance from the possession team's goal line). 
We account for play-context since these variables impact a team's designed play, which may result in a play with low or high STRAIN regardless of the pass rusher's role. 
For instance, a team may call a short pass that is intended to be thrown early which could limit the amount of STRAIN on a play. 
Or a team may need to throw a deep pass which would require more time and potentially create more STRAIN.
We do not account for time directly in the model due to the concern that the time it takes for a quarterback to throw the ball is itself a function of both the play call and pressure from pass rushers.
Thus, since we do not know the designed play call, we condition on the play context to adjust for play-level differences attributing to a pass rusher's STRAIN.

>>>>

In addition to the average STRAIN over all frames played, we can also calculate pass rusher $j$'s observed average STRAIN on a single play $i$ consisting of $T_i$ total frames,
\begin{equation*}
\overline{\text{STRAIN}}_{ij} = \frac{1}{T_i} \sum_{t=1}^{T_i} \widehat{\text{STRAIN}}_{ij}(t).
\end{equation*}
While this aggregated measure is a simple first step for pass rush evaluation, the observed average STRAIN on a single play is likely due to numerous factors.
Besides the pass rusher's ability, there is variability in the opposing strength of pass blockers across plays a pass rusher is involved, both at the individual and team levels.
Thus, we need to appropriately divide the credit of an observed average STRAIN across the different players and team involved, amongst other factors.

<!-- Additionally, based on previous work by PFF, the opposing QB plays an important role on the observed pressure in passing plays [@eager2019pressure]. We also suspect that pass rusher's STRAIN is affected by their defensive teammates as well.  -->

@@

In detail, we consider a normal distribution to shrink the random intercepts for each player and team toward their respective group means.
This is a useful property since we do not observe the same number of plays for each player. 
For the team effects, this provides us with the average defense and offense team-level effects on a pass rusher's STRAIN.
Due to the nested nature of players on teams, the individual pass rusher and blocker random intercepts reflect the respective player's effect relative to their team effects. 
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

@@

As for the fixed-effects about pass rusher $j$ in play $i$, we include a variety of features that likely contribute to variation in STRAIN.
First, we adjust for the position of both the pass rusher and nearest blocker to account for any positional effects.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
These are encoded as indicator variables with defensive ends and tackles as the reference levels for the pass rushers and blockers, respectively.
We also account for the number of pass blockers on the play, since teams may decide to employ a more protective scheme that could lower the observed STRAIN.
Finally, we control for play-context covariates with respect to the offensive team.
These include the current down (first, second, third, fourth, or two-point conversion), yards to go for a first down, and current yardline (i.e., distance from the possession team's goal line). 
We consider play-context information since these variables impact a team's designed play, which may result in a play with low or high STRAIN regardless of the pass rusher's role. 
For instance, a team may call a short pass that is intended to be thrown early which could limit the amount of STRAIN on a play. 
Or a team may need to throw a deep pass which would require more time and potentially create more STRAIN.
We do not account for time directly in the model due to the concern that the time it takes for a quarterback to throw the ball is itself a function of both the play call and pressure from pass rushers.
Thus, since we do not know the designed play call, we condition on the play context to adjust for play-level differences attributing to a pass rusher's STRAIN.

### 2023-07-18 paper/04-results.Rmd
updated bootstrap and more edits

As for play context, defensive pressure appears to increase when more yardage is required for the offense to reach a first down and for later plays within a set of down, when all other predictors are held constant.
In particular, every extra yard in the distance needed for a first down is associated with a 0.0491 increase (95% CI [0.0426, 0.0555]) in the average STRAIN.
Relative to first down situations, higher amount of defensive pressure seem to happen in plays that come after (second, third, and fourth downs).
In addition, we have insufficient evidence for a relationship between average STRAIN and the current yardline on the field for the offensive team.

@@

In addition to adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the varying intercept distributions for the top ten pass rushers for each position.
We obtain similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (defensive end), T.J. Watt (outside linebacker), Aaron Donald (defensive tackle), and Vita Vea (nose tackle) as the first-ranked pass rusher for their respective position.
Additionally, we observe different levels of variability across these top players, with only a subset of players with intercept ranges strictly above zero.
This is not necessarily surprising given the limited sample of data.

>>>>

As for play context, defensive pressure appears to increase when more yardage is required for the offense to reach a first down and for later plays within a set of down, when all other predictors are held constant.
In particular, every extra yard in the distance needed for a first down is associated with a 0.0491 increase (95% CI [0.0426, 0.0555]) in the average STRAIN.
Relative to first down situations, higher amount of pressure seem to happen in plays that come after (second, third, and fourth downs).
In addition, we have insufficient evidence for a relationship between average STRAIN and the current yardline on the field for the offensive team.

@@

Aside from adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the varying intercept distributions for the top ten pass rushers for each position.
We obtain similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (defensive end), T.J. Watt (outside linebacker), Aaron Donald (defensive tackle), and Vita Vea (nose tackle) as the first-ranked pass rusher for their respective position.
Additionally, we observe different levels of variability across these top defenders, with only a subset of players having intercept distributions strictly above zero.
This is not necessarily surprising given the limited sample of data.
